### 2024-09-26 paper/paper.qmd
Update paper.qmd

---
title: "Analysis of transit delays in the Toronto Transit Commission: 2024"
subtitle: "Insights from Jan 1 to Aug 31"
author: "Dong Jun Yoon"
thanks: "Code and data are available at: https://github.com/DJY1231/TTC-report-2024.git"
date: today
date-format: long
abstract: "This paper analyzes TTC delay data to understand patterns and causes of delays. It finds buses experience the most delays, especially on certain routes, while subways have frequent undocumented delays. Mechanical issues, operations, and diversions are most common delay causes. The analysis indicates the need to reduce bus delays through more vehicles or tactics like bus bridging on certain routes."
format: pdf
number-sections: true
bibliography: references.bib

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction
Having operated for over 100 years, the Toronto Transit Commission (TTC) has become one of the most visible and representative public transportation organizations in North America. It is well-known as the primary mode of public transportation in Toronto's daily life. Within the TTC, the bus network typically experiences higher daily ridership compared to streetcars, while subways accommodate the largest volume of passengers due to their central role in covering longer routes. Frequent riders often notice delays that arise on a daily basis. As delays become worse and services becoming increasingly inconsistent, it is important to examine the potential causes, allowing the TTC to identify and implement effective solutions.

This paper will identify the correlations of public transit delays across different types of transportation. The data utilized for analysis, including both records and census data, measurements, is in {#sec-data}. In {#sec-results}, I will talk about the predicted results derived from the data, this includes the result of the differences in delays between buses, streetcars, and subways. Lastly, {#sec-Discussion} contains more detailed results in a broader context and addresses weaknesses of this paper as well as the future explorations. This paper uses the programming language `R`[@citeR] . The analysis, the model and the visualizations use the following packages: `dplyr`[@dplyr],  `tidyverse`[@tidyverse], `ggplot2`[@ggplot], `knitr` [@knitr],  `here` [@here] , and `kableExtra` [@kableExtra],`modelsummary`[@model_summary], and `rstanarm`[@rstanarm].

# Data {#sec-data}
## Data Source
The data were gathered from Open Data Toronto's report on TTC delays. The datasets, published by the TTC, include the following: the first dataset presents delay statistics for subways in 2024 from January 1 to August 31; the second dataset details delay statistics for buses in 2024 from January 1 to August 31; and the third dataset provides delay statistics for streetcars in 2024 from January 1 to August 31.

# Data Analysis
The dataset provides information that offers insights of public transit delays across different types of transportation. The y-axis represents  The x-axis corresponds ,representing the total income for each year.

Additionally, data cleaning and analysis were conducted using the base statistical programming language R (R Core Team 2023), tidyverse (Wickham et al. 2019), ggplot2 (Wickham 2016), dplyr (Wickham et al. 2023), readr (Wickham, Hester, and Bryan 2023), tibble (Müller and Wickham 2023), janitor (Firke 2023), and knitr (Xie 2023) package were used.

# Overview of datasets
```{r}
install.packages("kableExtra")

library(kableExtra)
library(dplyr)
library(knitr)

data <- read.csv("ttc_bus_delay_data_2024.csv")

subset_data <- data %>%
  select(Date, Time, Day, Incident, Min.Delay, Min.Gap) %>% 
  head(10)

kable(subset_data, format = "pipe", align = 'l') %>%
  kable_styling(full_width = F, position = "center")

```
Table 1: Sample of bus delay statistics; the first 10 rows of TTC bus delay over time: Jan 1 to Aug 31, 2024.

```{r}
library(kableExtra)
library(dplyr)
library(knitr)

data <- read.csv("ttc_streetcar_delay_data_2024.csv")

subset_data <- data %>%
  select(Date, Time, Day, Incident, Min.Delay, Min.Gap) %>% 
  head(10)

kable(subset_data, format = "pipe", align = 'l') %>%
  kable_styling(full_width = F, position = "center")

```
Table 2: Sample of streetcar delay statistics; the first 10 rows of TTC subway delay over time: Jan 1 to Aug 31, 2024.

```{r}
library(kableExtra)
library(dplyr)
library(knitr)

data <- read.csv("ttc_subway_delay_data_2024.csv")

subset_data <- data %>%
  select(Date, Time, Day, Code, Min.Delay, Min.Gap) %>% 
  head(10)

kable(subset_data, format = "pipe", align = 'l') %>%
  kable_styling(full_width = F, position = "center")
```
Table 3: Sample of subway delay statistics; the first 10 rows of TTC subway delay over time: Jan 1 to Aug 31, 2024.

Table 1, Table 2, Table 3 include:

1. Date: The dataset shows the date when the delay occurred.
2. Time: The dataset contains information on the time the delay occurred.
3. Day: The dataset indicates the day of the week when the delay occurred.
4. Incident: The dataset displays the reason for the delay, such as a general delay, emergency services, or security issues. Incident is displayed in code for Table 3.
5. Min.Delay and Min.Gap: The dataset likely represents 'Min Delay' and 'Min Gap,' both measured in minutes. 'Min Delay' refers to the delay in the schedule for the next bus, while 'Min Gap' indicates the scheduled time, between the bus ahead and the following bus.

The total bus delay dataset contains 40,447 entries, while the subway delay dataset contains 17,517 entries, and streetcar delay dataset contains 9,932 entries. We can indicate that delays in the bus has occurred approximately 2.35 times more than the subway delay and 4.07 times more than the streetcar delay.

# Data Visualization
```{r}
# Load the required libraries
library(ggplot2)
library(dplyr)

# Read the data from both CSV files
bus_data <- read.csv("ttc_bus_delay_data_2024.csv")
streetcar_data <- read.csv("ttc_streetcar_delay_data_2024.csv")

# Combine the two data frames into one
combined_data <- bind_rows(bus_data, streetcar_data)

# Summarize the data by incident type (combining bus and streetcar delays)
incident_summary <- combined_data %>%
  group_by(Incident) %>%
  summarise(Total_Incidents = n()) %>%
  arrange(desc(Total_Incidents))

# Plot the total incidents for each incident type using ggplot2
ggplot(incident_summary, aes(x = reorder(Incident, -Total_Incidents), y = Total_Incidents)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Reasons of Bus and Streetcar Delay", 
       x = "Incident Type", 
       y = "Total Incidents") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for better readability

```
Figure 1: Distribution of Bus and Streetcar Delay Reasons

Figure 1 illustrates the main reasons for bus and streetcar delays. The primary cause of delays for bus and streetcars are mechanical issues, followed by operational issues, security concerns, diversions, and general delays. Other reasons can be observed in Figure 1, but I will not mention them all as they are not the main causes of delays.

```{r}
library(ggplot2)
library(dplyr)

subway_data <- read.csv("Subway_count_2024.csv")

subway_data_clean <- subway_data %>%
  filter(!is.na(Delay.Group) & !is.na(Count) & Count > 0)

# Summarize the data by Delay.Group
delay_group_summary <- subway_data_clean %>%
  group_by(Delay.Group) %>%
  summarise(Total_Count = sum(Count)) %>%  # Sum up the Count column for each Delay.Group
  arrange(desc(Total_Count))  # Arrange by total delays in descending order

# Plot the total incidents for each delay group using ggplot2
ggplot(delay_group_summary, aes(x = reorder(Delay.Group, -Total_Count), y = Total_Count)) +
  geom_bar(stat = "identity", fill = "tomato") +
  labs(title = "Total Delays by Delay Group (TTC Subway)", 
       x = "Delay Group", 
       y = "Total Delays") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for better readability

```

Figure 2: Distribution of Subway Delay Reasons

Figure 2 illustrates the main reasons for subway delays. The main cause of delays for the subway is patron, injury, incident, and mechanical issues. Other reasons can be observed in Figure 2, but I will not mention them all as they are not the main causes of delays.

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(lubridate)

# Define file paths (replace with actual paths)
file1 <- ("ttc_subway_delay_data_2024.csv")
file2 <- ("ttc_bus_delay_data_2024.csv")
file3 <- ("ttc_streetcar_delay_data_2024.csv")

data1 <- read.csv(file1)
data2 <- read.csv(file2)
data3 <- read.csv(file3)

# Inspect column names and structure
print(names(data1))
print(names(data2))
print(names(data3))

# Assuming the date column might be named 'Date'
data1$date <- as.Date(data1$Date, format = "%Y-%m-%d")
data2$date <- as.Date(data2$Date, format = "%Y-%m-%d")
data3$date <- as.Date(data3$Date, format = "%Y-%m-%d")

# Extract the day of the week for each file and count incidents
day_counts_file1 <- data1 %>%
  mutate(day_of_week = wday(date, label = TRUE)) %>%
  group_by(day_of_week) %>%
  summarise(total_count = n()) %>%
  mutate(file = "File 1")

day_counts_file2 <- data2 %>%
  mutate(day_of_week = wday(date, label = TRUE)) %>%
  group_by(day_of_week) %>%
  summarise(total_count = n()) %>%
  mutate(file = "File 2")

day_counts_file3 <- data3 %>%
  mutate(day_of_week = wday(date, label = TRUE)) %>%
  group_by(day_of_week) %>%
  summarise(total_count = n()) %>%
  mutate(file = "File 3")

# Combine the results into a single dataframe
combined_day_counts <- bind_rows(day_counts_file1, day_counts_file2, day_counts_file3)

# Plot the variation of delay incidents by day of the week for each file
ggplot(combined_day_counts, aes(x = day_of_week, y = total_count, color = file, group = file)) +
  geom_line() +
  geom_point() +
  labs(title = "Variation of Delay Incidents by Day of the Week",
       x = "Day of the Week",
       y = "Total Count",
       color = "File") +
  theme_minimal()
```
Figure 3:The trend of delays based on the day of the week

Figure 3 shows the trend of delays based on the day of the week. We can observe which day experiences the most delays and which day experiences the least. It also provides an overview of the delay trend throughout the week.

This information offers valuable insight for the TTC to prepare alternative solutions for delays. For instance, they could deploy more buses or trains on days with higher delays compared to days with fewer delays.

# Discussion

Relate your analysis to the themes mentioned in Bouie (2022), discussing the importance of remembering and analyzing historical events through data.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References
R Core Team. 2023. R: A Language and Environment for Statistical Computing. Vienna,
Austria: R Foundation for Statistical Computing. https://www.R-project.org/.